---
title: "Building phylogeny for Africa"
author: "Mathew Rees"
date: "15/03/2022"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up environment

Load packages

```{r, message=FALSE, error=FALSE, warning=FALSE}
library(tidyverse)
library(ape)
library(ggtree)
library(treeio)
library(tidytree)
library(V.PhyloMaker)
```

Load the data

```{r}
load("../2.Species_x_Site_Matrix/Merged.RData")

## There is an issue with families which gives me a dataframe with 50 extra rows

## Filter out collections with only genus name
sp2 <- merged %>% filter(! tax_sp_level == "")
## Select only variables of interest
sp2 <- sp2 %>% dplyr::select(family, genus, tax_sp_level)
(sp3 <- sp2 %>% group_by(species = tax_sp_level, genus, family) %>% summarise(count=n()))
``` 

We get a dataframe with 26,252 rows but only 26,202 species. Something is wrong.
There must be some duplicated names, perhaps with different families.

```{r}
## find out which are the duplicates
n_occur <- data.frame(table(sp3$species))
n_occur[n_occur$Freq > 1,]
#sp3[sp3$species %in% n_occur$Var1[n_occur$Freq > 1],]
```

Here are the culprits.

```{r}
## remove duplicated species
sp4 <- sp3[!duplicated(sp3$species),]

sp4$species.relative <- NA
sp4$genus.relative <- sp4$genus
sp4 <- sp4 %>% dplyr::select(!count)
sp4 <- as.data.frame(sp4)
str(sp4)
```

Ok so now we have our dataframe with 26,201 observations but the species factor has 26,202 levels. Not sure how this is possible but it seems to be ok with downstream analyses.

----------------------------------------------------

## Check how many names overlap

How many names in the merged dataset match those in V.Phylomaker?

```{r}
mega.tree <- V.PhyloMaker::GBOTB.extended
names <- mega.tree$tip.label

species_africa <- as.vector(sp4$species)
a1<- gsub(" ", "_", species_africa, fixed = TRUE) ## we want to add an "_" to our list from RAINBIO to match the style of the list of species in this phylogeny.

str(a1)
test <- names[names %in% a1]
str(test)

```

Out of 26,201 species, the V.Phylomaker `GBOTB.extended` tree has 6,284 so about 24%.

Note: after having ran the tree once, I get the following error message:

[1] "Taxonomic classification not consistent between sp.list and tree."
[1] "Note: 3 taxa fail to be binded to the tree,"
[1] "Hydnora_abyssinica"  "Hydnora_sinandevu"   "Jodrellia_fistulosa"

I will have to manually change some names. I've done this outside of R for practical reasons.

---------------------------------------------------

## Build the phylogeny

V.Phylomaker can make a tree by grafting tips onto the phylogeny according to the closest relative species or genus.

```{r}
# input example species list
example <- read.csv("V.Phylomaker_species_list_ecoregions.csv")
example <- example[!duplicated(example$species),]
str(example)

#generate the phylogeny presented in Figure 1b

rel <- bind.relative(sp.list=example, tree=GBOTB.extended, nodes=nodes.info.1)

tree.b <- phylo.maker(sp.list=rel$species.list, tree=rel$phylo, nodes=rel$nodes.info, scenarios="S3")
str(tree.b)

#write.tree(tree.b$scenario.3, "Africa.phylogeny.tre")
```

This is only one tree with random placement of the grafted tips. Ideally I should replicate this 100 times to get some uncertainty measures.
