---
title: "Community Phylogenetics"
author: "Mathew Rees"
date: "01/03/2022"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up the environment

First I load the libraries and data 

```{r message=FALSE, error=FALSE, warning=FALSE}
library(tidyverse)
library(ggtree)
library(treeio)
library(ape)
library(picante)
```

Now the data

```{r}
## Issue with tip labels of "Africa_phylogeny.tre"
tree <- read.tree("Africa_phylo_2.tre")
matrix.1 <- read.csv("Species_x_Ecoregions_Matrix.csv", row.names = 1, header = T)

## Make sure the names match up
names <- colnames(matrix)
names <- names %>% gsub(pattern = "\\ ", replacement = "_")
colnames(matrix.1) <- names
matrix.1 <- as.matrix(matrix.1)
```

Now match the two and calculate PD

```{r}
pd.result <- pd(matrix, tree, include.root = TRUE)
pd.result
```

Order them

```{r}
pd.result[order(pd.result$PD, decreasing = T),]
```

Seems like these two measures might be correlated

```{r}
library(ggpmisc)
ggplot(pd.result, aes(x=PD, y=SR)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq(formula = y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE)
```

Now look at measures of community dissimilarity. First make sure the names on the tips of the tree match the names in the species x ecoregions matrix.

```{r}
a<-tree$tip.label %>% sort()
b<-colnames(matrix.1) %>% sort()
identical(a,b)
```

Check which names are causing an issue

```{r}
a[!a %in% b]

b[!b %in% a]
```

Some weird NA is the tree and some strangely formated names in the matrix. Overall matrix names are better so I will replaces tree$tip.labels with the colnames of matrix.1

```{r}
tree$tip.label <- colnames(matrix.1)
```

Now try the ses.mpd

```{r}
memory.limit(size=50000)
phydist <- cophenetic(tree)
ses.mpd.result <- ses.mpd(matrix.1, phydist, null.model = "taxa.labels",
abundance.weighted = FALSE, runs = 99)
ses.mpd.result
```


```{r}
ses.mntd.result <- ses.mntd(matrix.1, phydist, null.model = "taxa.labels",
abundance.weighted = FALSE, runs = 99)
ses.mntd.result
```

