---
title: "Species x Site Matrix"
author: "Mathew Rees"
date: "15/03/2022"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the environment

```{r message=FALSE, error=FALSE, warning=FALSE}
library(tidyverse)
library(sp)
library(sf)
library(rgdal)
library(raster)
library(splancs)
library(ggsn)
library(maptools)
library(rspatial)
library(vegan)
library(reshape)
```

Load data
And remember Git has a 100MB file size limit
Can always use the command `git reset Head~`

```{r}
load("Merged.RData")

temp <- tempfile()
temp2 <- tempfile()

## Download ecoregions shapefile
#download.file("https://storage.googleapis.com/teow2016/Ecoregions2017.zip", temp)
#unzip(zipfile=temp, exdir = temp2)
#your_SHP_file<-list.files(temp2, pattern = ".shp$",full.names=TRUE)

##Read in the sahepfile
ecoregions <- st_read(your_SHP_file)
Africa1 <- ecoregions %>% filter(REALM == "Afrotropic")

rm(c(temp, temp2))
```


```{r}
sp<-merged

i <- sapply(sp, is.factor)                    ## here change the factors to characters
sp[i] <- lapply(sp[i], as.character)

coordinates(sp) <- c("decimalLongitude", "decimalLatitude")
proj4string(sp) <- CRS("+proj=longlat +datum=WGS84")

rrr <- as(Africa1, "Spatial")

ov <- over(sp, rrr)
head(ov)

v <- cbind(merged, ov)
table(v$ECO_NAME)
```

Plot

```{r}
spc <- tapply(v$tax_sp_level, v$ECO_NAME, function(x)length(unique(x)))
spc <- data.frame(ECO_NAME=names(spc), nspp = spc)
# merge with country SpatialPolygonsDataFrame
cn <- merge(rrr, spc, by='ECO_NAME')
print(spplot(cn, 'nspp', col.regions=rev(terrain.colors(25))))
```

Quick look at the relation between number of collections and species diversity.

```{r}
library(MASS)
library(ggpubr)
a <- table(v$ECO_NAME)
b <- as.data.frame(a)
c <- cbind(b, spc$nspp)
colnames(c) <- c("Ecoregions", "Number_of_records", "Number_of_species")
ggplot(c, aes(x=Number_of_records, y=Number_of_species)) + geom_point() + geom_smooth(method = "rlm") + theme_light() + stat_regline_equation(label.x = 2, label.y = 8000,aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")))
```

Ok so we already knew this in a sense but it shows that we need to collect more in under-explored areas! Areas with less diversity will be less collected because they are less interesting. The two outliers on the far right are the Sahelian Acacia savanna and the West Sudanian savanna.

## Reshape the data

Code from here: https://www.r-bloggers.com/2012/07/r-for-ecologists-creating-a-site-x-species-matrix/

Use the reshape package

```{r}
v$value <- 1 ## create a value column for the cast function
site.sp.quad <- cast(v, formula = ECO_NAME ~ tax_sp_level, value = 'value')
site.sp.quad <- as.data.frame(site.sp.quad)
## Remove the last row which is just NA
site.sp.quad <- site.sp.quad[-95,]
matrix <- site.sp.quad
names<- matrix$ECO_NAME
## Remove first column and turn it into rownames
matrix <- matrix[,-1]
rownames(matrix) <- names
## Turn all values into presence/absence data with 0 or 1
matrix[matrix>1] <- 1

#write.csv(matrix, "Species_x_Ecoregions_Matrix.csv")
```

Try this